Large-scale feature selection with Gaussian mixture models for the classification of high dimensional remote sensing images

### 2017-02-06 paper.tex
update

% \documentclass[journal,peerreview,onecolumn]{IEEEtran}
\documentclass[journal,10pt]{IEEEtran}
\usepackage{etex}
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\graphicspath{{./Fig/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{url}
\usepackage{listings}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tabularx}
\usepackage{empheq}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{minted}\definecolor{bg}{rgb}{0.95,0.95,0.95}
\newcommand\MyBox[2]{
  \fbox{\lower0.75cm
    \vbox to 1.2cm{\vfil
      \hbox to 1.2cm{\hfil\parbox{1.4cm}{#1\\#2}\hfil}
      \vfil}%
  }%
}

@@

% correct bad hyphenation here
% \hyphenation{op-tical net-works semi-conduc-tor}
\newcommand{\rev}[1]{\textcolor{magenta}{\emph{#1}}}

@@

% \tableofcontents
% \clearpage
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
  A  large  scale  feature  selection wrapper  is  discussed  for  the
  classification  of high  dimensional  remote  sensing. An  efficient
  implementation is proposed based on intrinsic properties of Gaussian
  mixtures models and  block matrix.  The criterion  function is split
  into two parts  : one that is  updated to test each  feature and one
  that needs to be updated only once per feature selection. This split
  saved  a  lot  of  computation  for each  test.   The  algorithm  is
  implemented in  \texttt{C++} and integrated into  the Orfeo Toolbox.
  It has been compared to  other classification algorithms on two high
  dimension remote sensing images. Results show that the approach
  provides good classification accuracies with low computation time.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
% \begin{IEEEkeywords}
% remote sensing, hyperspectral imaging, feature selection, gaussian mixture model, fast computing.
% \end{IEEEkeywords}

@@

% Among the various dimension reduction techniques, feature selection demonstrates several appealing for practical situation. The objective of feature selection is to find a subset of the original features with a minimum loss according the objective. Thus,

% retrieve the most compact representation of the data with minimum loss of information \cite{Guyon:2006:FEF:1208773}. By deleting the redundant information, feature selection allows to reduce dimension and thus avoid statistical issue, increase algorithm speed and limit storage requirements. Moreover, it adds several beneficial side effects. It improves data understanding by identify where is the most relevant information and it is then possible to save resources when organizing a new data acquisition.

@@

% The method proposed in this paper is actually a wrapper method which has been optimized in order to accelerate model learning and testing which is the weak point of this family.

@@

% Finally, tests are conducted to compare different variations of the method and to compare also to other classifiers (GMM, Random Forest, k-nearest-neighbor) available in the Orfeo Toolbox .

@@

    For mixture models, it is assumed  that a given sample $\mathbf{x}$ is
    the realization  of a  random vector which  distribution is  a mixture
    (convex     combination)     of      several     class     conditioned
    distributions~\cite{Fraley00model-basedclustering}:
    \begin{align}
        p(\mathbf{x}) = \sum_{c=1}^{C} \pi_c f_c(\mathbf{x}|\theta)
    \end{align}
    where $\pi_c$ is  the prior, \emph{i.e.}, the  proportion of class
    $c$  and  $f_c$  a  parametric density  function  controlled by
    $\theta$.

@@

    \begin{figure}[!t]
        \centering
        \begin{tikzpicture}
          \begin{axis}[ymode=log,xmin=0,xmax=200,ymin=0,width=0.8\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = $d$,ylabel = $\eta_c$,small]
            \addplot+[very thick,mark=none,smooth,domain=0:200,samples=200] (\x,{\x*(\x+3)/2+1});
            \end{axis}
        \end{tikzpicture}
        \caption{Number of parameters $\eta_c$ per class in function of dimension $d$: $\eta_c=d(d+3)/2+1$.\label{fig:nb-param}}
    \end{figure}

@@

    % Criterion functions aim to evaluate either a rate of correct classification or the separability/similarity of class distributions. These functions are used to estimate which sets of variables are the best to represent data, to assure the separability of the classes and to perform classification. The choice of a criterion function is the choice of a way to compare sets of variables.

@@

        \hspace{0pt} \\

@@

        \vspace{10 mm}

        % According to \cite{bruzzone2009novel}, it is interesting to notice that the KL divergence increases quadratically with respect to the distance between the mean vectors of the class distributions whereas the measures of correct classification asymptotically tends to one when distributions are perfectly separable. On the contrary, the JM distance tends to saturate as the measures of correct classification.

@@

        % Then, the inverse of  the covariance matrix
        % $(\boldsymbol{\Sigma}^{(k)})^{-1}$,   the   quadratical   term
        % $(\mathbf{x}^{(k)})^t         (\boldsymbol{\Sigma}^{(k)})^{-1}
        % \mathbf{x}^{(k)}$        and        the        log-determinant
        % $\log   |\boldsymbol{\Sigma}^{(k)}|$  can   be  expressed   in
        % function of terms of the $(k-1)^{th}$ iteration.

@@

        % In the case of a backward step in SFFS algorithm, the formula is inverted in order to compute $(\boldsymbol{\Sigma}^{(k-1)})^{-1}$ knowing $(\boldsymbol{\Sigma}^{(k)})^{-1}$. The update rule becomes (proof in Appendix~\ref{app:proof-update})
        % \begin{prop}[Backward update rule for inverse of covariance matrix]
        % \label{eq:update-inv-back}
        %     \begin{equation*}
        %         (\boldsymbol{\Sigma}^{(k-1)})^{-1} = \underbrace{\mathbf{A}}_{\substack{\text{computed once}\\ \text{per selection step}}} - \underbrace{\alpha \mathbf{v} \mathbf{v}^t}_{\substack{\text{computed for} \\ \text{each augmented set}}}
        %     \end{equation*}
        % \end{prop}

@@

      \begin{prop}[Jeffries-Matusita distance~(\ref{eq:jm})]
      \begin{align}
        \label{eq:jm-update}
            \text{B}_{cc^\prime} = &\underbrace{\frac{1}{4} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{2} \log \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_c^{(k-1)}| |\boldsymbol{\Sigma}_{c^\prime}^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 + \frac{1}{2} \log \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_c \alpha_{c^\prime}}} \right)}_{\substack{\text{computed for each augmented set}}},
      \end{align}
    \end{prop}
    \noindent where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha_c$ and $\mathbf{v}_c$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}_c$.

@@

    \begin{figure}[!t]
        \centering
        \begin{tabular}{c}
            \includegraphics[width=0.7\columnwidth]{Fig/aisa.png} \\
            {\bfseries{(a)}} \\
            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt.png} \\
            {\bfseries{(b)}} \\
        \end{tabular}
        \caption{Aisa dataset: {\bfseries{(a)}} colored composition of the image (R: 634nm, G: 519nm, B: 477nm), {\bfseries{(b)}} ground-truth.\label{fig:aisa}}
    \end{figure}

@@

    \begin{figure}[!t]
        \centering
        \begin{tikzpicture}
            \begin{axis}[ymin=0.2,ymax=0.9,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
                \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl_aisa.txt};
                \draw[thick,red] (axis cs:18,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:18,\pgfkeysvalueof{/pgfplots/ymax});
                \draw[thick,black] (axis cs:24,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:24,\pgfkeysvalueof{/pgfplots/ymax});
            \end{axis};
            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl.txt};
            %     \draw[thick,red] (axis cs:15,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:15,\pgfkeysvalueof{/pgfplots/ymax});
            %     \draw[thick,black] (axis cs:47,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:47,\pgfkeysvalueof{/pgfplots/ymax});
            % \end{axis};
            % \begin{axis}[ymin=0.2,ymax=0.75,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa}]
            %     \addplot[thick,black] table[x=nb,y=crit] {criterion_50000spl.txt};
            %     \draw[thick,red] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
            %     \draw[thick,black] (axis cs:31,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:31,\pgfkeysvalueof{/pgfplots/ymax});
            % \end{axis};
        \end{tikzpicture}
        \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Aisa dataset with 500 samples by class. Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.\label{fig:crit-evol}}
    \end{figure}

    % Then, a small processing is made to select the number of variables used for prediction. The evolution of the criterion function is retrieved from the model file and the discrete derivative of the criterion \emph{DJ} is computed and normalized by its maximum. The subset of variables used for classification $\Omega^{*}$ is the first subset for which this value is lower than a given threshold \emph{th}, equal  to 0.1\% in this experiment,
    % \begin{equation*}
    %     \Omega^{*} = \Omega_{\min \{k|DJ_k<th\}}
    % \end{equation*}
    % with $DJ_k = \frac{J(\Omega_k) - J(\Omega_{k-1})}{\max_k (J(\Omega_k) - J(\Omega_{k-1}))}$.

@@

    When creating training and validation  sets, special care is taken
    to assure that training samples are picked out from distinct areas
    than test  samples.  The  polygons of the  reference are  split in
    smaller polygons and then 50\%  of the polygons are taken randomly
    for training and the remaining 50\% for validation.  An example of
    training      and     validation      set     is      shown     in
    Figure~\ref{fig:set-aisa}.  From the  training  polygons, a  given
    number of samples  were selected to build the  training set, while
    all  the pixels  from the  validation polygons  were used  for the
    validation.  Moreover 20  random trials were run  with a different
    training  set (different  polygons).  Table~\ref{tab:aisa-otbsimu}
    presents  the results  of the  experiment with  mean and  standard
    deviation  of  the  Kappa  coefficient  over  the  20  trials  and
    Table~\ref{tab:aisa-otbsimu-time}  the   corresponding  processing
    time. Bold values corresponds to best results. In Table~\ref{tab:aisa-otbsimu}, when several bold scores appears for the same experiment, it means that the scores has been assessed as equivalent with a Wilcoxon rank-sum test \cite{mann1947test}. Additionally, Figure~\ref{fig:meanNbVar-aisa} summarizes the mean of the number of selected variables for each variation of the GMM classifier with selection.

    \begin{figure}[!t]
        \centering
        \begin{tabular}{c}
            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt_train.png} \\
            {\bfseries{(a)}} \\
            \includegraphics[width=0.7\columnwidth]{Fig/aisa_gt_test.png} \\
            {\bfseries{(b)}} \\
        \end{tabular}
        \caption{Aisa dataset: {\bfseries{(a)}} training polygons of first trial, {\bfseries{(b)}} test polygons of first trial.\label{fig:set-aisa}}
    \end{figure}

@@

    % this one is result without changing the number of selected variables
    % \begin{table}[!t]
    %     \centering
    %     \caption{Average classification accuracy 20 trials (standard deviation in parenthesis).\label{tab:aisa-otbsimu}}
    %     \begin{tabular}{lccc}\toprule
    %          & \multicolumn{3}{c}{\bfseries Cohen's kappa} \\ \cmidrule{2-4}
    %         \# samples by class & 250 & 500 & 1000 \\ \midrule

    %         GMM SFS kappa & {\bfseries 0.684 (0.025)} & {\bfseries 0.698 (0.027)} & {\bfseries 0.711 (0.028)} \\
    %         GMM SFS JM &    {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
    %         GMM SFFS JM &   {\bfseries 0.680 (0.028)} & {\bfseries 0.700 (0.028)} & {\bfseries 0.710 (0.030)} \\
    %         GMM ridge &     0.611 (0.040) & 0.620 (0.036) & 0.642 (0.034) \\
    %         KNN &           0.551 (0.035) & 0.563 (0.033) & 0.574 (0.030) \\
    %         Random Forest & 0.645 (0.026) & 0.673 (0.023) & 0.693 (0.023) \\
    %         \bottomrule
    %     \end{tabular}
    % \end{table}

    \begin{figure}
      \centering
      \begin{tikzpicture}
        \begin{axis}[ylabel=$\bar{n}_s$,ymax=25,
          symbolic x coords={0,250,500,1000,1050},
          enlargelimits=0.15,legend pos=outer north east,xtick=data,legend cell align=left,
          ybar,
          bar width=5pt,grid,xlabel=Number of training samples per class,small,area legend]
          \addplot coordinates {(250,15.9) (500,21.85) (1000,26.65)};
          \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
          \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
          %%\addplot[red,very thick,sharp plot,update limits=false]coordinates {(0,252) (1050,252)};
          \legend{{\footnotesize SFS kappa},{\footnotesize SFS JM},{\footnotesize SFFS JM}}
        \end{axis}
      \end{tikzpicture}
      \caption{Mean number $\bar{n}_s$ of selected features for the different variation of selection methods for Aisa data set. The original number of features is 252.}
      \label{fig:meanNbVar-aisa}
    \end{figure}

@@

@@

    The processing times of the three standard classifiers suffer from the increase of training samples. For the GMM classifier with ridge, the selection of regularization parameter is more costly with more samples because of the classification rate estimation needed. For the KNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file. Finally, for the Random Forest classifier, the trees tends to be deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.
    % The computation of the kappa seems  to suffer from  a lack of  parallelization   which could be  improved. ->  plus tard dans la conclusion.

@@

    % It can be guessed that the fact that the Gaussian hypothesis is unlikely to be verified explains the difficulty to extract the finer information available in a larger dataset. The JM distance has stressed in Table~\ref{tab:crit} has a strong Gaussian hypothesis and the same hypotheses could also explain the drop of the classification rate of GMM classifiers with JM distance as criterion.

@@

% Finally, an improvement could be made to increase the stability of the selected features. For example, with hyperspectral data, a selection of continuous intervals and not band is a possible solution which has already been explored in \cite{serpico2007extraction}.

% The python and C++ code are available freely for download: \url{https://www.orfeo-toolbox.org/external-projects/}, .

\appendices
\section{Proof of update rules}
\label{app:proof-update}
    % \subsubsection{Proposition~(\ref{eq:update-inv-back})}
    %     \begin{proof}
    %         \begin{align*}
    %             \mathbf{A} - \alpha \mathbf{v} \mathbf{v}^t
    %             &= (\boldsymbol{\Sigma}^{(k-1)})^{-1} + \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1} \\
    %             &~~~- \alpha (- \frac{1}{\alpha} (\boldsymbol{\Sigma}^{(k-1)})^{-1} \mathbf{u}) (- \frac{1}{\alpha} \mathbf{u}^t (\boldsymbol{\Sigma}^{(k-1)})^{-1})) \\
    %             &= (\boldsymbol{\Sigma}^{(k-1)})^{-1}
    %         \end{align*}
    %     \end{proof}

@@

    % \begin{align*}
    %     \log &\left(|\boldsymbol{\Sigma}_c^{(k)}|\right) =
    %     \log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{u}_c      \\
    %             \mathbf{u}_c^t          & \sigma^{(k)}_c \\
    %         \end{array}\bigg]
    %     \right|\right) \\
    %     =&\log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
    %             \mathbf{u}_c^t          & 1 \\
    %         \end{array}\bigg]
    %         \bigg[\begin{array}{cc}
    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
    %             \mathbf{0}_{k-1,1}^t          & \sigma^{(k)}_c - \mathbf{u}_c^t(\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c\\
    %         \end{array}\bigg]
    %     \right|\right) \\
    %     =&\log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \boldsymbol{\Sigma}_c^{(k-1)} & \mathbf{0}_{k-1,1}     \\
    %             \mathbf{u}_c^t          & 1 \\
    %         \end{array}\bigg]
    %     \right|\right)
    %     +\log \left(\left|
    %         \bigg[\begin{array}{cc}
    %             \mathbf{I}_{k-1} & (\boldsymbol{\Sigma}_c^{(k-1)})^{-1} \mathbf{u}_c      \\
    %             \mathbf{0}_{k-1,1}^t          & \alpha_c \\
    %         \end{array}\bigg]
    %     \right|\right) \\
    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| |1| \right) + \log \left( |\mathbf{I}_{k-1}| |\alpha_c| \right)\\
    %     =&\log \left( |\boldsymbol{\Sigma}_c^{(k-1)}| \right) + \log ( \alpha_c )
    % \end{align*}
    % \noindent where $\mathbf{I}_{k-1}$ is the $(k-1) \times (k-1)$ identity matrix and $\mathbf{0}_{k-1,1}$ is the $(k-1) \times 1$ null vector.
  From eq.~(\ref{eq:cov}) and standard results for the determinant of block matrix~\cite[Chapter 9]{IMM2012-03274} we have immediately:
  \begin{align*}
    \log \left(|\boldsymbol{\Sigma}_c^{(k)}|\right)  & = \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\sigma^{(k)}_c  - \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}  \mathbf{u}_c\right)\\
    &= \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\alpha_c\right)
  \end{align*}

@@

% \begin{IEEEbiography}{Michael Shell}
% Biography text here.
% \end{IEEEbiography}

@@

%\begin{figure}[!t]

>>>>

% \documentclass[journal,peerreview,onecolumn]{IEEEtran}
\documentclass[journal,10pt]{IEEEtran}
\usepackage{etex}
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\graphicspath{{./Fig/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}
\usepackage[caption=false,font=footnotesize]{subfig}
% \usepackage{fixltx2e} % it trigers a warning. It may be useless ?
\usepackage{url}
\usepackage{listings}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{tabularx}
\usepackage{empheq}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{minted}\definecolor{bg}{rgb}{0.95,0.95,0.95}
\newcommand\MyBox[2]{
  \fbox{\lower0.75cm
    \vbox to 1.2cm{\vfil
      \hbox to 1.2cm{\hfil\parbox{1.4cm}{#1\\#2}\hfil}
      \vfil}%
  }%
}

@@

% correct bad hyphenation here
% \hyphenation{op-tical net-works semi-conduc-tor}
% \newcommand{\rev}[1]{\textcolor{black}{\emph{#1}}}
\newcommand{\rev}[1]{\textcolor{black}{#1}}

@@

% \tableofcontents
% \clearpage
\begin{abstract}
  A  large  scale  feature  selection wrapper  is  discussed  for  the
  classification  of high  dimensional  remote  sensing. An  efficient
  implementation is proposed based on intrinsic properties of Gaussian
  mixtures models and  block matrix.  The criterion  function is split
  into two parts  : one that is  updated to test each  feature and one
  that needs to be updated only once per feature selection. This split
  saved  a  lot  of  computation  for each  test.   The  algorithm  is
  implemented in  \texttt{C++} and integrated into  the Orfeo Toolbox.
  It has been compared to  other classification algorithms on two high
  dimension remote sensing images. Results show that the approach
  provides good classification accuracies with low computation time.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
remote sensing, hyperspectral imaging, feature selection, gaussian mixture model, fast computing.
\end{IEEEkeywords}

@@

@@

@@

@@

    For mixture models, it is assumed  that a given sample $\mathbf{x}$ is
    the realization  of a  random vector which  distribution is  a mixture
    (convex     combination)     of      several     class     conditioned
    distributions~\cite{Fraley00model-basedclustering}:
    \begin{align}
        p(\mathbf{x}) = \sum_{c=1}^{C} \pi_c f_c(\mathbf{x}|\theta),
    \end{align}
    where $\pi_c$ is  the prior, \emph{i.e.}, the  proportion of class
    $c$  and  $f_c$  a  parametric density  function  controlled by
    $\theta$.

@@

    \begin{figure}[!t]
        \centering
        \begin{tikzpicture}
          \begin{axis}[ymode=log,xmin=0,xmax=200,width=0.8\columnwidth,grid,axis x line=left ,axis y line=left, tick align=outside,xlabel = $d$,ylabel = $\eta_c$,small]
            \addplot+[very thick,mark=none,smooth,domain=0:200,samples=200] (\x,{\x*(\x+3)/2+1});
            \end{axis}
        \end{tikzpicture}
        \caption{Number of parameters $\eta_c$ per class in function of dimension $d$: $\eta_c=d(d+3)/2+1$.\label{fig:nb-param}}
    \end{figure}

@@

@@

@@

@@

@@

@@

      \begin{prop}[Jeffries-Matusita distance~(\ref{eq:jm})]
      \begin{align}
        \label{eq:jm-update}
            &\text{B}_{cc^\prime} = \underbrace{\frac{1}{4} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})^t ( \boldsymbol{\tilde{\Sigma}}^{(k-1)} )^{-1} (\boldsymbol{\mu}_c^{(k-1)} - \boldsymbol{\mu}_{c^\prime}^{(k-1)})}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{2} \log \left( \frac{|\boldsymbol{\tilde{\Sigma}}^{(k-1)}|}{\sqrt{|\boldsymbol{\Sigma}_c^{(k-1)}| |\boldsymbol{\Sigma}_{c^\prime}^{(k-1)}|}} \right)}_{\substack{\text{computed once per selection step}}} \nonumber \\
            &+ \underbrace{\frac{1}{4} \tilde{\alpha} ( \left[\begin{array}{cc} \mathbf{\tilde{v}}^t & \frac{1}{\tilde{\alpha}} \end{array}\right] (\boldsymbol{\mu}_c^{(k)} - \boldsymbol{\mu}_{c^\prime}^{(k)}) )^2 + \frac{1}{2} \log \left( \frac{\tilde{\alpha}}{\sqrt{\alpha_c \alpha_{c^\prime}}} \right)}_{\substack{\text{computed for each augmented set}}},
      \end{align}
    \end{prop}
    \noindent where $\boldsymbol{\tilde{\Sigma}} = \boldsymbol{\Sigma}_c + \boldsymbol{\Sigma}_{c^\prime}$ and $\tilde{\alpha}$ and $\mathbf{\tilde{v}}$ are defined as $\alpha_c$ and $\mathbf{v}_c$ but using $\boldsymbol{\tilde{\Sigma}}$ instead of $\boldsymbol{\Sigma}_c$.

@@

    \begin{figure}[!t]
        \centering
        \begin{tabular}{c}
            \includegraphics[width=0.6\columnwidth]{Fig/aisa.png} \\
            {\bfseries{(a)}} \\
            \includegraphics[width=0.6\columnwidth]{Fig/aisa_gt.png} \\
            {\bfseries{(b)}} \\
        \end{tabular}
        \caption{Aisa dataset: {\bfseries{(a)}} colored composition of the image (R: 634nm, G: 519nm, B: 477nm), {\bfseries{(b)}} ground-truth.\label{fig:aisa}}
    \end{figure}

@@

    \begin{figure}[!t]
        \centering
        \begin{tikzpicture}
            \begin{axis}[ymin=0.2,ymax=0.9,grid,axis x line=left,axis y line=left,xlabel={\# variables},ylabel={Cohen's kappa},small]
                \addplot[thick,black] table[x=nb,y=crit] {criterion_1000spl_aisa.txt};
                \draw[thick,red] (axis cs:18,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:18,\pgfkeysvalueof{/pgfplots/ymax});
                \draw[thick,black] (axis cs:24,\pgfkeysvalueof{/pgfplots/ymin}) -- (axis cs:24,\pgfkeysvalueof{/pgfplots/ymax});
            \end{axis};
        \end{tikzpicture}
        \caption{Criterion evolution (kappa) in function of the number of selected variables for first trial with Aisa dataset with 500 samples by class. Red vertical line is the retained number of variables and black vertical line is the maximum of the criterion.\label{fig:crit-evol}}
    \end{figure}

@@

    When creating training and validation  sets, special care is taken
    to assure that training samples are picked out from distinct areas
    than test  samples.  The  polygons of the  reference are  split in
    smaller polygons and then 50\%  of the polygons are taken randomly
    for training and the remaining 50\% for validation.  An example of
    training      and     validation      set     is      shown     in
    Figure~\ref{fig:set-aisa}.  From the  training  polygons, a  given
    number of samples  were selected to build the  training set, while
    all  the pixels  from the  validation polygons  were used  for the
    validation.  Moreover 20  random trials were run  with a different
    training  set (different  polygons).  Table~\ref{tab:aisa-otbsimu}
    presents  the results  of the  experiment with  mean and  standard
    deviation  of  the  Kappa  coefficient  over  the  20  trials  and
    Table~\ref{tab:aisa-otbsimu-time}  the   corresponding  processing
    time. Bold values corresponds to best results. In Table~\ref{tab:aisa-otbsimu}, when several bold scores appears for the same experiment, it means that the scores has been assessed as equivalent with a Wilcoxon rank-sum test \cite{mann1947test}. Additionally, \ref{tab:aisa-otbsimu-time} summarizes the mean of the number of selected variables for each variation of the GMM classifier with selection.

    \begin{figure}[!t]
        \centering
        \begin{tabular}{c}
            \includegraphics[width=0.6\columnwidth]{Fig/aisa_gt_train.png} \\
            {\bfseries{(a)}} \\
            \includegraphics[width=0.6\columnwidth]{Fig/aisa_gt_test.png} \\
            {\bfseries{(b)}} \\
        \end{tabular}
        \caption{Aisa dataset: {\bfseries{(a)}} training polygons of first trial, {\bfseries{(b)}} test polygons of first trial.\label{fig:set-aisa}}
    \end{figure}

@@

@@

    % \begin{figure}
    %   \centering
    %   \begin{tikzpicture}
    %     \begin{axis}[ylabel=$\bar{n}_s$,ymax=25,
    %       symbolic x coords={0,250,500,1000,1050},
    %       enlargelimits=0.15,legend pos=outer north east,xtick=data,legend cell align=left,
    %       ybar,
    %       bar width=5pt,grid,xlabel=Number of training samples per class,small,area legend]
    %       \addplot coordinates {(250,15.9) (500,21.85) (1000,26.65)};
    %       \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
    %       \addplot coordinates {(250,11.95) (500,12) (1000,12.05)};
    %       \legend{{\footnotesize SFS kappa},{\footnotesize SFS JM},{\footnotesize SFFS JM}}
    %     \end{axis}
    %   \end{tikzpicture}
    %   \caption{Mean number $\bar{n}_s$ of selected features for the different variation of selection methods for Aisa data set. The original number of features is 252.}
    %   \label{fig:meanNbVar-aisa}
    % \end{figure}

@@

    The processing times of the three standard classifiers suffer from the increase of training samples. For the GMM classifier with ridge, the selection of regularization parameter is more costly with more samples because of the classification rate estimation needed. For the KNN classifier, the model stores all the training samples and the prediction implies to compute the distance to all the training samples which explains the increase of the processing time and additionally of the size of the model file. Finally, for the Random Forest classifier, the trees tends to be deeper in order to capture the additional information available with more samples and that explains the increase of the processing time.

@@

@@

\appendices
\section{Proof of update rules}
\label{app:proof-update}

@@

  From eq.~(\ref{eq:cov}) and standard results for the determinant of block matrix~\cite[Chapter 9]{IMM2012-03274} we have immediately:
  \begin{align*}
    \log \left(|\boldsymbol{\Sigma}_c^{(k)}|\right)  & = \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\sigma^{(k)}_c  - \mathbf{u}_c^t (\boldsymbol{\Sigma}_c^{(k-1)})^{-1}  \mathbf{u}_c\right)\\
    &= \log\left(|\boldsymbol{\Sigma}_c^{(k-1)}|\right)\log\left(\alpha_c\right)
  \end{align*}

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Fig/lagrange.jpg}}]{Adrien Lagrange} received an Engineering degree in Robotics and Embedded Systems from ENSTA ParisTech, Paris (Palaiseau), France, and the M.Sc. degree in Machine Learning from the Paris Saclay University, both in 2016.

He is currently a Ph.D. student at the National Polytechnic Institute of Toulouse, within the Signal and Communications Group of the IRIT Laboratory. He is working on the subject of multi-resolution learning for hierarchical analysis of hyperspectral and hypertemporal images under the supervision of Nicolas Dobigeon and Mathieu Fauvel. His research interests are remote sensing, statistical modeling, image processing.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fauvel.pdf}}]{Mathieu Fauvel}
 graduated in  electrical engineering  from the Grenoble  Institut of
 Technology (Grenoble  INP), Grenoble, France, in  2004.  He  received
 the M.S and  Ph.D.  degrees in image and signal  processing from  the
 Grenoble INP  in 2004  and 2007,  respectively.  In  2007, he  was a
 teaching assistant  in Grenoble  INP. From  2008 to  2010, he  was a
 postdoctoral research associate with the MISTIS Team of the  National
 Institute  for Research  in  Computer Science  and Control  (INRIA).
 Since 2011,  Dr.  Fauvel  has been an  Associate Professor  with  the
 National Polytechnic  Institute of  Toulouse (ENSAT -  University of
 Toulouse) within  the DYNAFOR lab  (University of Toulouse  -  INRA).
 His  research interests  are  remote sensing,  data fusion,  pattern
 recognition, multicomponent signal and image processing. From  2013 to 2016,
 he was the president of the French Chapter of the GRSS. He is  now in charge of the European GRSS Chapters activities.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Fig/manuel}}]{Manuel Grizonnet} received the Mathematical modeling, Vision, Graphics and
  Simulation Engineer degree from the \'{E}cole Nationale Sup\'{e}rieure d'Informatique
  et de Math\'{e}matiques Appliqu\'{e}es de Grenoble, France, in 2007. From 2007 to 2009
  he worked for the BRGM (French geological survey) in Niamey, Niger where he was a
  Systems and Geological Information System (GIS) engineer. He is currently with
  the Centre National d'\'{E}tudes Spatiales (French Space Agency), Toulouse,
  France, where he is developing image processing algorithms and software for
  the exploitation of Earth observation images.
\end{IEEEbiography}

@@
